### 2009-03-02 doc/storage_system/arc-storage-documentation.tex
update of the storage documentation

In the first prototype release there is one client tool called \verb!arc_storage_cli!, which is written in Python, and only need a basic Python installation to run. It is capable of communicating with a given Bartender service, and uploading and downloading TURLs via HTTP.

The methods can be listed with:
\begin{verbatim}
$ arc_storage_cli
Usage:
  arc_storage_cli <method> [<arguments>]
Supported methods: stat, make[Collection], unmake[Collection], list, move,
    put[File], get[File], del[File]
\end{verbatim}
Without arguments, each method prints its own help:
\begin{verbatim}
$ arc_storage_cli move
Usage: move <sourceLN> <targetLN>
\end{verbatim}

Uploading, downloading and stat files:

\begin{verbatim}
$ cat testfile 
This is a testfile.
$ arc_storage_cli put testfile /tmp/
- The size of the file is 20 bytes
- The md5 checksum of the file is 9a9dffa22d227afe0f1959f936993a80
- ARC_BARTENDER_URL environment variable not found, using http://localhost:60000/Bartender
- Calling the Bartender's putFile method...
- done in 0.08 seconds.
- Got transfer URL: http://localhost:60000/hopi/d15900f5-34ee-4bba-bb10-73d60d1c0d75
- Uploading from 'testfile'
    to 'http://localhost:60000/hopi/d15900f5-34ee-4bba-bb10-73d60d1c0d75' with http...
Uploading 20 bytes... data sent, waiting... done.
- done in 0.0042 seconds.
'testfile' (20 bytes) uploaded as '/tmp/testfile'.
$ arc_storage_cli stat /tmp/testfile
- ARC_BARTENDER_URL environment variable not found, using http://localhost:60000/Bartender
- Calling the Bartender's stat method...
- done in 0.05 seconds.
'/tmp/testfile': found
  states
    checksumType: md5
    neededReplicas: 1
    size: 20
    checksum: 9a9dffa22d227afe0f1959f936993a80
  timestamps
    created: 1210232135.57
  parents
    51e12fab-fd3d-43ec-9bc5-17041da3f0b2/testfile: parent
  locations
    http://localhost:60000/Shepherd fc0d3d99-6406-4c43-b2eb-c7ec6d6ab7fe: alive
  entry
    type: file
$ arc_storage_cli get /tmp/testfile newfile
- ARC_BARTENDER_URL environment variable not found, using http://localhost:60000/Bartender
- Calling the Bartender's getFile method...
- done in 0.05 seconds.
- Got transfer URL: http://localhost:60000/hopi/dab911d0-110f-468e-b0c3-627af6e3af31
- Downloading from 'http://localhost:60000/hopi/dab911d0-110f-468e-b0c3-627af6e3af31'
    to 'newfile' with http...
Downloading 20 bytes... done.
- done in 0.0035 seconds.
'/tmp/testfile' (20 bytes) downloaded as 'newfile'.
$ cat newfile 
This is a testfile.
\end{verbatim}

You can find more examples in the SVN\footnote{\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/storage/README}}.

There are plans to create more sophisticated CLI and GUI tools and to create a FUSE\footnote{Filesystem in Userspace, \url{http://fuse.sourceforge.net/}} module and Windows Shell Extensions\footnote{\url{http://msdn.microsoft.com/en-us/magazine/cc188741.aspx}} to be able to mount the ARC storage namespace into the local filesystem namespace, and use it with the commands of the operating system.

@@

\section{Integrating third-party storage solutions} % (fold)
\label{sec:integrating_third_party_storage_solutions}

To integrate existing files on a third-party storage system to our namespace thus make them accessible through the interface of the ARC storage, we need a service which provides a common interface to the Bartender services, and hides the details of accessing the different third-party storages. It should translate the method calls, the \emph{get}s, \emph{put}s and \emph{remove}s and the \emph{ACL}\footnote{Access Control Lists, policies} modifications, and try to create transform metadata according to the data model of the ARC storage (Section~\ref{sec:data_model}). The files here would be referenced by a path which is local in the namespace of the third-party storage. The interface could be something like this:
\begin{description}
    \item[get(getRequest)] returns list of (\emph{requestID}, \emph{getResponseData})

    very similar to the \emph{get} method of the Shepherd

    \item[put(putRequest)] returns a list of (\emph{requestID}, \emph{putResponseData})

    very similar to the \emph{put} method of the Shepherd

    \item[delete(deleteRequest)] returns a list of (\emph{requestID}, \emph{success})

    very similar to the \emph{delete} method of the Shepherd

    \item[stat(statRequest)] returns a list of (\emph{requestID}, \emph{statResponse})

    very similar to the \emph{stat} method of the Shepherd

    \item[list(listRequest)] returns \emph{listResponse}

    very similar to the \emph{list} method of the Bartender

    \item[move(moveRequest)] returns a list of (\emph{requestID}, \emph{status})

    very similar to the \emph{move} method of the Bartender

\end{description}
This interface has some methods similar to the Shepherd and some other methods similar to the Bartender.
% section integrating_third_party_storage_solutions (end)

>>>>

\subsection{Prototype CLI tool} % (fold)
\label{sub:prototype_cli_tool}

In the first prototype release there is one client tool called \verb!arc_storage_cli!, which is written in Python, and only need a basic Python installation to run. It is capable of communicating with a given Bartender service, and uploading and downloading TURLs via HTTP. However in order to use proxy certificates or to transfer files with GridFTP protocol, you need to have the ARC client libraries installed.

The credentials of the user and the URL of the Bartender can be configured either with an XML file located at \verb!~/.arc/client.xml! or with environment variables.

Example for the \verb!client.xml! file:

\begin{verbatim}
    <ArcConfig>
      <ProxyPath>/Users/zsombor/Development/arc/proxy</ProxyPath>
      <CACertificatesDir>/Users/zsombor/Development/arc/certs/CA</CACertificatesDir>
      <BartenderURL>https://localhost:60000/Bartender</BartenderURL>
    </ArcConfig>
\end{verbatim}

Example for the environment variables:

\begin{verbatim}
    export ARC_BARTENDER_URL=https://localhost:60000/Bartender
    export ARC_KEY_FILE=/Users/zsombor/Development/arc/certs/userkey-john.pem
    export ARC_CERT_FILE=/Users/zsombor/Development/arc/certs/usercert-john.pem 
    export ARC_CA_DIR=/Users/zsombor/Development/arc/certs/CA    
\end{verbatim}

You can specify the number of needed replicas for new file with the \verb!ARC_NEEDED_REPLICAS! environment variable.

The \verb!arc_storage_cli! has its built-in help. The methods can be listed with:

\begin{verbatim}
$ arc_storage_cli
Usage:
  arc_storage_cli <method> [<arguments>]
  Supported methods: stat, make[Collection], unmake[Collection], list, move,
    put[File], get[File], del[File], pol[icy], unlink,
    credentialsDelegation, removeCredentials
\end{verbatim}
Without arguments, each method prints its own help:
\begin{verbatim}
$ arc_storage_cli move
Usage: move <sourceLN> <targetLN>
\end{verbatim}

Here is an example of uploading, downloading and stat files:

\begin{verbatim}
$ cat testfile 
This is a testfile.
$ arc_storage_cli put testfile /tmp/
- The size of the file is 20 bytes
- The md5 checksum of the file is 9a9dffa22d227afe0f1959f936993a80
- Calling the Bartender's putFile method...
- done in 0.08 seconds.
- Got transfer URL: http://localhost:60000/hopi/d15900f5-34ee-4bba-bb10-73d60d1c0d75
- Uploading from 'testfile'
    to 'http://localhost:60000/hopi/d15900f5-34ee-4bba-bb10-73d60d1c0d75' with http...
Uploading 20 bytes... data sent, waiting... done.
- done in 0.0042 seconds.
'testfile' (20 bytes) uploaded as '/tmp/testfile'.
$ arc_storage_cli stat /tmp/testfile
- Calling the Bartender's stat method...
- done in 0.05 seconds.
'/tmp/testfile': found
  states
    checksumType: md5
    neededReplicas: 1
    size: 20
    checksum: 9a9dffa22d227afe0f1959f936993a80
  timestamps
    created: 1210232135.57
  parents
    51e12fab-fd3d-43ec-9bc5-17041da3f0b2/testfile: parent
  locations
    http://localhost:60000/Shepherd fc0d3d99-6406-4c43-b2eb-c7ec6d6ab7fe: alive
  entry
    type: file
$ arc_storage_cli get /tmp/testfile newfile
- Calling the Bartender's getFile method...
- done in 0.05 seconds.
- Got transfer URL: http://localhost:60000/hopi/dab911d0-110f-468e-b0c3-627af6e3af31
- Downloading from 'http://localhost:60000/hopi/dab911d0-110f-468e-b0c3-627af6e3af31'
    to 'newfile' with http...
Downloading 20 bytes... done.
- done in 0.0035 seconds.
'/tmp/testfile' (20 bytes) downloaded as 'newfile'.
$ cat newfile 
This is a testfile.
\end{verbatim}

% subsection prototype_cli_tool (end)

\subsection{FUSE module} % (fold)
\label{sub:fuse_module}

TODO: what is the FUSE module, how does it work, how can we use it

% subsection fuse_module (end)

@@

\section{Grid integration} % (fold)
\label{sec:grid_integration}

TODO: what is the ARC DMC, how does it work

% section grid_integration (end)
